Kernel Square-Loss Exemplar Machines

### 2016-11-15 main_cvpr_review.tex
sections 5 and 6
version J-1

% Highlighting
\usepackage{soul}
%\usepackage[usenames,dvipsnames]{xcolor}
\newcommand{\hlc}[2][yellow]{{\sethlcolor{#1}\hl{#2}}}
\newcommand{\RAF}[1]{\hlc[yellow]{(RR:) #1}}
\newcommand{\JZ}[1]{\hlc[pink]{(JZ:) #1}}
\newcommand{\PP}[1]{\hlc[green]{PP: #1}}
\newcommand{\un}[1]{\underline{#1}}

@@

\maketitle
%\thispagestyle{empty}
%%%%%%%%% ABSTRACT
\begin{abstract}
This paper proposes an extension of the exemplar SVM (ESVM) as a novel approach to feature encoding for feature retrieval applications.
%This paper proposes a new feature encoding pipeline, based on the success of the exemplar SVM (ESVM) \cite{ZePe15} and linear discriminant analysis (LDA) as whitening \cite{HMR12}.
We first show that, by replacing the hinge-loss by the square-loss in the ESVM cost function, equivalent results in image retrieval can be obtained at a fraction of the computational cost. 
We call this model square-loss exemplar machine, or SLEM. 
We also introduce a kernelized SLEM which can be implemented efficiently through a low-rank matrix decomposition and
%benefits from the same computational advantages but 
displays improved performance. 
Both SLEM variants exploit the fact that the negative examples are fixed for all the positives in the training set, so most of the SLEM computational complexity is relegated to an offline process independent of the positive examples. 
Our experiments establish the performance and computational advantages of our approach using a large array of base feature representations and standard image retrieval datasets.
\end{abstract}

@@

%% RR: uncomment for final version
%% \subsection*{Acknowledgments} This work was supported by the ERC grant VideoWorld.

{\small
\bibliographystyle{ieee} 
\bibliography{sup,jz}
}
\end{document}

>>>>

% Highlighting
\usepackage{soul}
%\usepackage[usenames,dvipsnames]{xcolor}
\newcommand{\hlc}[2][yellow]{{\sethlcolor{#1}\hl{#2}}}
\newcommand{\RAF}[1]{\hlc[yellow]{(RR:) #1}}
\newcommand{\JZ}[1]{\hlc[pink]{(JZ:) #1}}
\newcommand{\PP}[1]{\hlc[green]{PP: #1}}
\newcommand{\un}[1]{\underline{#1}}
\def\dul#1{\underline{\underline{#1}}}

@@

\maketitle
%\thispagestyle{empty}
%%%%%%%%% ABSTRACT
\begin{abstract}
Zepeda and P\'erez \cite{ZePe15} have recently demonstrated the promise of the exemplar SVM (ESVM) as a feature encoder for image retrieval. This paper extends this approach in several directions:
%This paper proposes an extension of the exemplar SVM (ESVM) as a novel approach to feature encoding for feature retrieval applications.
%This paper proposes a new feature encoding pipeline, based on the success of the exemplar SVM (ESVM) \cite{ZePe15} and linear discriminant analysis (LDA) as whitening \cite{HMR12}.
We first show that replacing the hinge loss by the square-loss in the ESVM cost function significantly reduces encoding time with negligible effect on
accuracy.
%equivalent results in image retrieval can be obtained at a fraction of the computational cost. 
We call this model square-loss exemplar machine, or SLEM. 
We then introduce a kernelized SLEM which can be implemented efficiently through a low-rank matrix decomposition and
%benefits from the same computational advantages but 
displays improved performance. 
Both SLEM variants exploit the fact that the negative examples are fixed for all the positives in the training set, so most of the SLEM computational complexity is relegated to an offline process independent of the positive examples. 
Our experiments establish the performance and computational advantages of our approach using a large array of base feature representations and standard image retrieval datasets.
\end{abstract}

@@

%% RR: uncomment for final version
%% \subsection*{Acknowledgments} This work was supported by the ERC grant VideoWorld.
%\newpage
{\small
\bibliographystyle{ieee} 
\bibliography{sup,jz}
}
\end{document}

### 2016-11-15 sec_SLEM.tex
version J-1

\section{Square loss exemplar machine}\label{lsesvm}
In this section, we revisit the exemplar SVM model proposed in \cite{Efros11} as an instance of a more general family of classifiers. Then, we introduce the square loss exemplar machine (SLEM) by solving a simple variant of this model and study its properties.
\subsection{Exemplar classifiers} \label{esvm}
We are given base features in $\RR^d$ at training time, one positive example $x_0$ in $\RR^d$ and a set of negative examples $X = [x_1, x_2,...,x_n]$ in $\RR^{d\times n}$, each column of $X$ representing one example by a vector in $\RR^d$. 
We are also given a loss function $l:\{-1, 1\}\times \RR\rightarrow\RR^+$. Learning an exemplar classifier from these examples amounts to minimizing the function 
\begin{equation}
J(\omega, \nu) = \theta \ l(1, \omega^Tx_0+\nu) +\dfrac{1}{n}\sumi l(-1, \omega^T x_i+\nu)+\dfrac{\lambda}{2}\|\omega\|^2, 
\label{eq:first}
\end{equation}
w.r.t. $\omega$ in $\RR^d$ and $\nu$ in $\RR$. In Eq. (\ref{eq:first}), $\lambda$ and $\theta$ are respectively a regularization parameter on $\omega$ and a positive scalar adjusting the weight of the positive exemplar.
%\footnote{We could also acknowledge a parameter $\theta$ other than $\frac{1}{n}$ as a regularization to the error of each negative example. But this parameters seems to be less important to cross-validate. Also, setting $\theta=\frac{1}{n}$ simplify Equation (\ref{omega:solution}) and allows the use of Woodbury identity.}

Given a cost $l$, we define the correspondent \emph{exemplar classifiers} of $x_0$ with respect to $X$ are the weights $\omega^\star(x_0,X)$ that minimizes the loss function $J$:
\begin{equation}
\big(\omega^\star, \nu^\star\big) = \underset{(\omega,\nu)\in\RR^d\times\RR}{\argmin} \ J(\omega, \nu). \footnote{Depending on the loss function $l$, $\nu^\star(x_0,X)$ may be not unique. \PP{But, what about $w^{\star}$?}}
%\big(\omega^\star(x_0, X), \nu^\star(x_0, X)\big) = \underset{(\omega,\nu)\in\RR^d\times\RR}{\argmin} \ J(\omega, \nu). 
\label{omega:first}
\end{equation}
%To shorten the notation, we refer to these solutions as $\omega^\star$ and $\nu^\star$ when their arguments are implicit.

%\cite{Efros11} introduced ensembles of exemplar classifiers for object detection, but single exemplar classifiers have been used as replacement of the original feature $x_0$ for image classification and retrieval \cite{ZePe15} and detection \cite{GMPD12}. \RAF{Maybe this should be in the introduction instead.}
The exemplar SVM~\cite{Efros11,Efros12} is an instance of this model where $l$ is the hinge loss. 
%Applications of ESVM use the hinge loss function, which guarantees that $J$ is a convex function. 
The solution of (\ref{omega:first}) can be thus found by stochastic gradient descent~\cite{bottou10} individually for each positive sample. The next subsection show how to calculate all exemplar classifiers simultaneously by changing the loss function.

\subsection{Square loss function}\label{slem:intro}
Now, let us study the same learning problem for the square loss function $l(y,\hat{y}) = \frac{1}{2}(y-\hat{y})^2$. As in the case of the hinge loss, Equation (\ref{eq:first}) is a convex problem. 
However it is now a ridge regression problem, whose unique solution can be found in closed form as
\begin{align}
\begin{cases}
\vspace{3 mm}
\omega^\star &= \dfrac{2\theta}{\theta+1}U^{-1}(x_0-\mu), \\
%\vspace{3 mm}
\nu^\star &= \dfrac{\theta-1}{\theta+1}-\dfrac{1}{\theta+1}(\theta x_0+\mu)^T\omega^\star,
\end{cases}
\label{omega:solution}
\end{align}
%\begin{align}
%\nu^\star &= \dfrac{\theta-1}{\theta+1}-\dfrac{1}{\theta+1}(\theta x_0+\mu)^T\omega^\star, \label{eq:nustar}\\
%\omega^\star &= \dfrac{2\theta}{\theta+1}U^{-1}(x_0-\mu) \label{eq:wstar}
%\end{align}
where:
\begin{align}
\begin{cases}
\mu = &\frac{1}{n}\sum_{i=1}^n x_i,\\ \\
U = &\frac{1}{n}XX^T-\mu\mu^T\\
&+\frac{\theta}{\theta+1}(x_0-\mu)(x_0-\mu)^T+\lambda\mathrm{Id}_d. \label{eq:U}
\end{cases}
\end{align}
Equation (\ref{omega:solution}) shows how to solve (\ref{eq:first}) simultaneously for all individual exemplars.
\cite{lssvm}
%Replacing the hinge loss by the square loss offers a more compact solution, but not necessarily more efficient.

@@

Note that the positive sample weight $\theta$ does not influence the direction of the optimal vector $\omega^\star$, only its norm. This means that if search and ranking are based on the normalized feature $\frac{1}{\|\omega^\star\|}\omega^\star$, \eg using cosine similarity, $\theta$ does not influence the matching score of the SLEM vectors of two different images. This sets SLEM appart from ESVM which requires this parameter to be calibrated \cite{Efros11,ZePe15}. We can thus set $\theta=\frac{1}{n}$ for the remaining of this work.
%The normalized exemplar classifier $\omega^\star$ is therefore a linear transformation of $x_0-\mu$.
%Indeed, if $\omega$ and $\omega'$ are the $d$-dimensional SLEM vectors of $x$ and $x'$, respectively, their cosine similarity reads:
%we can denote $s$ the matching score scalar function defined in $\RR^d\times \RR^d$, which is given by
%\begin{equation}
%s(\omega, \omega') = \dfrac{\omega^T \omega'}{\|\omega\| \|\omega'\|} = \frac{(x-\mu)^TA^{-2}(x'-\mu)}{\|A^{-1}(x-\mu)\|\|A^{-1}(x'-\mu)\|},\label{match:score}
%\end{equation}
%which does not depend on the value of $\theta$. 
%This means that the weight of the positive sample can be fixed at any positive value without changing matching scores.

@@

It is interesting to note the relationship between SLEM and the classical linear discriminant analisys (LDA).
Let us return to Eq. (\ref{eq:first}) an suppose that we have multiple positive samples. It can be shown that in this case, the corresponding linear classifier of Eq. (\ref{eq:first}) for the square loss is also given by
(\ref{omega:solution}), where $x_0$ denotes this time the center of mass
of the positive samples {\em if} these samples have the {\em same} covariance matrix $\Sigma$ as the negative samples $X$.

This equal-covariance assumption is of course quite restrictive, and probably unrealistic in general. It is interesting to note, however, that this is exactly the assumption made by linear discriminant analysis. As shown in~\cite{Hastie2009} for example, LDA can be seen as a (non-regularized) linear classifier with decision function $\omega^T_{LDA} z+ \nu_{LDA}$, where $z$ is a sample in
$\RR^d$, and
\begin{equation}
\left\{\begin{array}{l}
\displaystyle \omega_{LDA}=\Sigma^{-1}(x_0-\mu),\\
\displaystyle \nu_{LDA}=-\frac{1}{2}(x_0+\mu)^T \omega_{LDA}.
\end{array}\right.
\label{eq:lda}
\end{equation}

This shows that SLEM is a generalized version of LDA: Indeed, when $\lambda=0$ in SLEM (\ie, no regularization), $A = \Sigma$, the vectors $\omega_{LDA}$ of Eq. (\ref{eq:lda}) and $\omega^\star$ of Eq. (\ref{Wood:omega}) have the same direction.
%This observation has been equally made by \cite{Koba15}.
Many interesting properties of LDA have been used recently for classification tasks \cite{GMPD12,HMR12} and, more recently, for image retrieval with convolutional neural network activations \cite{babenko15}.
With our simple generalization of LDA, we hope to obtain superior results.

>>>>

\section{The square-loss exemplar machine}\label{lsesvm}
In this section, we revisit the exemplar SVM model proposed in \cite{Efros11} as an instance of a more general family of classifiers. Then, we introduce the square loss exemplar machine (SLEM) by solving a simple variant of this model and study its properties.
\subsection{Exemplar classifiers} \label{esvm}
We are given base features in $\RR^d$ at training time, one positive example $x_0$ in $\RR^d$ and a set of negative examples $X = [x_1, x_2,...,x_n]$ in $\RR^{d\times n}$, each column of $X$ representing one example by a vector in $\RR^d$. 
We are also given a loss function $l:\{-1, 1\}\times \RR\rightarrow\RR^+$. Learning an exemplar classifier from these examples amounts to minimizing the function 
\begin{equation}
J(\omega, \nu) = \theta \ l(1, \omega^Tx_0+\nu) +\dfrac{1}{n}\sumi l(-1, \omega^T x_i+\nu)+\dfrac{\lambda}{2}\|\omega\|^2, 
\label{eq:first}
\end{equation}
w.r.t. $\omega$ in $\RR^d$ and $\nu$ in $\RR$. In Eq. (\ref{eq:first}), $\lambda$ and $\theta$ are respectively a regularization parameter on $\omega$ and a positive scalar adjusting the weight of the positive exemplar.
%\footnote{We could also acknowledge a parameter $\theta$ other than $\frac{1}{n}$ as a regularization to the error of each negative example. But this parameters seems to be less important to cross-validate. Also, setting $\theta=\frac{1}{n}$ simplify Equation (\ref{omega:solution}) and allows the use of Woodbury identity.}

Given a cost $l$, we define the correspondent \emph{exemplar classifiers} of $x_0$ with respect to $X$ are the weights $\omega^\star(x_0,X)$ that minimizes the loss function $J$:
\begin{equation}
\big(\omega^\star, \nu^\star\big) = \underset{(\omega,\nu)\in\RR^d\times\RR}{\argmin} \ J(\omega, \nu). \footnote{Depending on the loss function $l$, $\nu^\star(x_0,X)$ may be not unique.}
%\big(\omega^\star(x_0, X), \nu^\star(x_0, X)\big) = \underset{(\omega,\nu)\in\RR^d\times\RR}{\argmin} \ J(\omega, \nu). 
\label{omega:first}
\end{equation}
%To shorten the notation, we refer to these solutions as $\omega^\star$ and $\nu^\star$ when their arguments are implicit.

%\cite{Efros11} introduced ensembles of exemplar classifiers for object detection, but single exemplar classifiers have been used as replacement of the original feature $x_0$ for image classification and retrieval \cite{ZePe15} and detection \cite{GMPD12}. \RAF{Maybe this should be in the introduction instead.}
The exemplar SVM~\cite{Efros11,Efros12} is an instance of this model where $l$ is the hinge loss. 
%Applications of ESVM use the hinge loss function, which guarantees that $J$ is a convex function. 
The solution of Eq. (\ref{omega:first}) can thus be found by stochastic gradient descent~\cite{bottou10} individually for each positive sample. The next section shows how to calculate all exemplar classifiers simultaneously by changing the loss function.

\subsection{The square-loss}\label{slem:intro}
Now, let us study the same learning problem for the square-loss function $l(y,\hat{y}) = \frac{1}{2}(y-\hat{y})^2$. As in the case of the hinge loss, the minimization of Eq. (\ref{eq:first}) is a convex problem. 
However it is now a ridge regression problem, whose unique solution can be found in closed form as
\begin{align}
\begin{cases}
\vspace{3 mm}
\omega^\star &= \dfrac{2\theta}{\theta+1}U^{-1}(x_0-\mu), \\
%\vspace{3 mm}
\nu^\star &= \dfrac{\theta-1}{\theta+1}-\dfrac{1}{\theta+1}(\theta x_0+\mu)^T\omega^\star,
\end{cases}
\label{omega:solution}
\end{align}
%\begin{align}
%\nu^\star &= \dfrac{\theta-1}{\theta+1}-\dfrac{1}{\theta+1}(\theta x_0+\mu)^T\omega^\star, \label{eq:nustar}\\
%\omega^\star &= \dfrac{2\theta}{\theta+1}U^{-1}(x_0-\mu) \label{eq:wstar}
%\end{align}
where:
\begin{align}
\begin{cases}
\mu = &\frac{1}{n}\sum_{i=1}^n x_i,\\ \\
U = &\frac{1}{n}XX^T-\mu\mu^T\\
&+\frac{\theta}{\theta+1}(x_0-\mu)(x_0-\mu)^T+\lambda\mathrm{Id}_d. \label{eq:U}
\end{cases}
\end{align}
%Replacing the hinge loss by the square loss offers a more compact solution, but not necessarily more efficient.

@@

Note that the positive sample weight $\theta$ does not influence the direction of the optimal vector $\omega^\star$, only its norm. This means that if search and ranking are based on the normalized feature $\frac{1}{\|\omega^\star\|}\omega^\star$, \eg using cosine similarity, $\theta$ does not influence the matching score of the SLEM vectors of two different images. This sets SLEM appart from ESVM which requires this parameter to be calibrated \cite{Efros11,ZePe15}. We can thus set $\theta=\frac{1}{n}$ for the remaining of this work.

%Equation (\ref{Wood:omega}) also shows how to solve (\ref{eq:first}) in closed form and, hence, simultaneously for all individual exemplars by solving a linear system~\cite{lssvm}. If $X_p$ is the matrix where every column represents one positive sample, we obtain their respective exemplar classifier by solving the system $A\Omega = \bar{X}_p$ in $\Omega$. Here, $\bar{X}_p$ denotes the $X_p$ centralized w.r.t. the negatives' mean. It 
%The normalized exemplar classifier $\omega^\star$ is therefore a linear transformation of $x_0-\mu$.
%Indeed, if $\omega$ and $\omega'$ are the $d$-dimensional SLEM vectors of $x$ and $x'$, respectively, their cosine similarity reads:
%we can denote $s$ the matching score scalar function defined in $\RR^d\times \RR^d$, which is given by
%\begin{equation}
%s(\omega, \omega') = \dfrac{\omega^T \omega'}{\|\omega\| \|\omega'\|} = \frac{(x-\mu)^TA^{-2}(x'-\mu)}{\|A^{-1}(x-\mu)\|\|A^{-1}(x'-\mu)\|},\label{match:score}
%\end{equation}
%which does not depend on the value of $\theta$. 
%This means that the weight of the positive sample can be fixed at any positive value without changing matching scores.

@@

It is interesting to note the relationship between SLEM and the classical linear discriminant analysis (LDA).
Let us return to Eq. (\ref{eq:first}) an suppose that we have multiple positive samples. It can be shown that in this case, the corresponding linear classifier of Eq. (\ref{eq:first}) for the square-loss is also given by
(\ref{omega:solution}), where $x_0$ denotes this time the center of mass
of the positive samples {\em if} these samples have the {\em same} covariance matrix $\Sigma$ as the negative samples $X$.

This equal-covariance assumption is of course quite restrictive, and probably unrealistic in general. It is interesting to note, however, that this is exactly the assumption made by linear discriminant analysis. As shown in~\cite{Hastie2009} for example, LDA can be seen as a (non-regularized) linear classifier with decision function $\omega^T_{LDA} z+ \nu_{LDA}$, where $z$ is a sample in
$\RR^d$, and
\begin{equation}
\left\{\begin{array}{l}
\displaystyle \omega_{LDA}=\Sigma^{-1}(x_0-\mu),\\
\displaystyle \nu_{LDA}=-\frac{1}{2}(x_0+\mu)^T \omega_{LDA}.
\end{array}\right
\label{eq:lda}
\end{equation}

This shows that SLEM is a generalized version of LDA: Indeed, taking $\lambda=0$ (\ie no regularization), $A = \Sigma$ and the vectors $\omega_{LDA}$ of Eq. (\ref{eq:lda}) and $\omega^\star$ of Eq. (\ref{Wood:omega}) have the same direction, reducing SLEM to LDA.
%This observation has been equally made by \cite{Koba15}.
Many interesting properties of LDA have been used recently for classification tasks \cite{GMPD12,HMR12} and, more recently, for image retrieval with convolutional neural network activations \cite{babenko15}.
With our simple generalization of LDA, we hope to obtain superior results.

### 2016-11-15 sec_eff_imp.tex
version J-1

\input{figure_LR.tex}
When compared to the linear square-loss classifier of Section \ref{slem:intro}, one drawback of a kernelized approach is that the dimension of our problem grows with the size $n$ of the negative samples.
The offline factorization $B$ of $K$ demands $O(nr)$ storage and at best $O(nr^2)$ time. This factorization can be obtained in different ways, depending on the aspect we wish to optimize. Efficiency in time and efficiency in storage are the two most common of these aspects. In this section we propose three different decompositions of $K$, to be applied depending if we want to minimize time or storage.
%on which of these two efficiencies we hope to achieve.

@@

\subsection{Time-efficient implementation}
The Cholesky decomposition is the most used factorization of positive-semidefinite matrices in kernel-based learning \cite{BaJo02,BaJo05,FiSc01}. 
The computation of the factor $B$ depends on a permutation $\pi(n) =\{i_1,i_2,\dots,i_n\}$ of $\{1,2,\dots, n\}$, called \textit{pivots}. The $t$-th column of $B$ is calculated iteratively as:

\begin{align}
\begin{cases}
\vspace{3 mm}
B(i_t, t) = \left(K(i_t,i_t)-\sum_{m=1}^{t-1} B(i_{m},m)  \right)^{\frac{1}{2}},\\
\vspace{3 mm}
B(i_j,t) = 0,\\
B(i_j, t) = \frac{1}{B(i_t,t)}(K(i_j,i_t)  -\sum_{m=1}^{t-1}B(i_j,m)B(i_t,m)), \end{cases}\label{icd:algo}
\end{align}
for all $t<j\le n$. There are two possible choices of Cholesky decomposition. The incomplete Cholesky decomposition (ICD) stops its algorithm after $r$ steps and has complexity $O(nr^2)$. 
To minimize the residue $tr(K-BB^T)$, the $t$-th pivot is calculated at the end of the $t$-th interaction. 
In the other hand, the complete Cholesky decomposition (CCD) consists of a full-rank decomposition, \textit{i.e.} the rank $r$ of $B$ is equal to $n$ and it has complexity $O(n^3)$. 
In this case, the pivots are irrelevant to the final residue (which is zero) and can be assumed to be $i_t=t$. %The computation of pivots makes CCD more efficient for similar residues.
We compare CCD and ICD performances in Fig. \ref{fig:residue}. We see that for small $r$, ICD is indeed faster. As we increase the value fo $r$, ICD becomes slower than CCD due to the pivot calculation but its residue decreases. We further compare both methods in Section \ref{time-scale}.

We make sure $K$ is positive-definite by adding $\epsilon$ to its diagonal, where $\epsilon=\min(0,-\lambda_{min})$ and $\lambda_{min}$ is the smallest eigenvalue of $K$. Therefore, from the equation $BB^T=K+\epsilon\mathrm{Id}_n$, $B$ has rank $n$ and we can apply CCD.

\subsection{Storage-efficient implementation}\label{low-rank} %/Low-rank decomposition} 
Let us now consider instead the problem of improving storage efficiency. As discussed in section \ref{simi_score}, for each positive exemplar we store its original encoder plus a $2r$ vector, so we aim to decompose $K$ at a small rank $r$.
For small values of $r$, a kernel PCA (KPCA) factorization guarantees a better approximation of $K$ than the Cholesky decomposition at a higher time cost. Indeed, as illustrated in Figure \ref{fig:residue}, a kernel PCA gives smaller residues than ICD at small rank. We plot the ratio $tr(K-BB^T)/tr(K)$ when we vary the number of columns $r$ of $B$, and we see a faster convergence for the kernel PCA decomposition when $r$ goes to $n$. However, at these small values of $r$, KPCA is more time consuming than both ICD and CCD. This time cost comes from applying a singular values decomposition of K to obtain $B$:
\begin{equation}
    K = WDW^T; \ b_i = \sqrt{d_{ii}}w_i, \label{svd}
\end{equation}
where $w_i$ denotes the $i$-th column of $W$ and $d_{ii}$ the $i$-th diagonal element of $D$. Decomposing $K$ from Equation~(\ref{svd}) has complexity $O(n^2r)$ in time. \hlc{Francis, I need a reference for this complexity.}

ICD may be the fastest decomposition for small ranks, but it is outperformed by KPCA; for higher ranks, ICD is the slowest alternative and its performance approximates CCD's performance. We therefore use KPCA for our storage-efficient. %\hlc{RR: However, ICD is still the only $O(nr^2)$ in time, which could be useful for bigger values of $n$. I'm considering showing the experiments done in the ECCV supmat to show why ICD is obsolete.}

%\begin{figure}
%    \centering
%    \includegraphics[width=0.5\textwidth]{trace_residual.jpg}
%    \caption{Residual error of decompositions varying by its rank.}
%    \label{fig:residue}
%\end{figure}

>>>>

\input{figure_LR.tex}
When compared to the linear square-loss classifier of Section \ref{slem:intro}, one drawback of the kernelized approach is that the dimension of our problem grows with the size $n$ of the negative samples. The offline factorization $B$ of $K$ demands $O(nr)$ storage and at best $O(nr^2)$ time. 
This factorization can be obtained in two fundamental ways: full-rank and low-rank decomposition. In this section we propose three different decompositions of $K$, to be applied depending on the compromise between accuracy and storage we went to achieve.
%on which of these two efficiencies we hope to achieve.

@@

\subsection{Full-rank decomposition}
The complete Cholesky decomposition (CCD) is the most used factorization of positive-definite matrices in kernel-based learning due to its time efficiency \cite{BaJo02}. We use it as our default decomposition.
We make sure $K$ is positive-definite by adding $\epsilon$ to its diagonal, where $\epsilon=\min(0,-\lambda_{min})$ and $\lambda_{min}$ is the smallest eigenvalue of $K$. Therefore, from the equation $BB^T=K+\epsilon\mathrm{Id}_n$, $B$ has rank $n$ and can be obtained by CCD.

\subsection{Low-rank decomposition}
Let us now turn to the problem of improving storage efficiency. One of the major requisites of a global image descriptor for large scale retrieval is to minimize the storage.
As discussed in Section \ref{simi_score}, for each positive exemplar we store its base feature plus a $2r$ vector. Hence, we aim to decompose $K$ at a small rank $r$.
We consider the problem of finding a factor $B$ of fixed rank $r$ such that the residue $\mathrm{tr}(K-BB^T)/\mathrm{tr}(K)$ is small. \hlc{I wanted to write "factor $B$ of fixed rank $r$ that minimizes the residue $\mathrm{tr}(K-BB^T)/\mathrm{tr}(K)$", but I don't want to propose it as an optimization problem. The "such that the residue is small" is not perfect, but it is the best I could find. Any proposition?}

The incomplete Cholesky decomposition (ICD) is widely used in machine learning~\cite{BaJo02,BaJo05,FiSc01}. Its algorithm is similar to CCD, and it greedily chooses which column of $K$ to add to the decomposition based on the gain in approximation error for all non-added columns. This process is called \emph{pivot selection}~\cite{BaJo05}. We stop the algorithm $r$ steps, obtaining the factor $B$ in time $O(nr^2)$.

The kernel PCA (KPCA) proposes that, instead of a greedy addition of columns, we choose the $r$ first eigenvectors. This is done by doing a truncated singular value decomposition of $K$:
\begin{equation}
    K = WDW^T; \ b_i = \sqrt{d_{ii}}w_i, \label{svd}
\end{equation}
where $w_i$ denotes the $i$-th column of $W$ and $d_{ii}$ the $i$-th diagonal element of $D$. Decomposing $K$ from Equation~(\ref{svd}) has complexity $O(n^2r)$ in time.~\cite{golub2012matrix}

We plot in Fig. \ref{fig:residue} the ratio $\mathrm{tr}(K-BB^T)/\mathrm{tr}(K)$ when we vary the number of columns $r$ of $B$ and we see a faster convergence for the kernel PCA decomposition when $r$ goes to $n$.
\hlc{I could not finish the discussion in this section, I need some sleep...}

%The incomplete Cholesky decomposition (ICD) is the most used factorization of positive-semidefinite matrices in kernel-based learning due to its time efficiency \cite{BaJo02,BaJo05,FiSc01}. The computation of the factor $B$ depends on a permutation $\pi(n) =\{i_1,i_2,\dots,i_n\}$ of $\{1,2,\dots, n\}$, called \textit{pivots}. The $t$-th column of $B$ is calculated iteratively as:

%\begin{align}
%\begin{cases}
%\vspace{3 mm}
%B(i_t, t) = \left(K(i_t,i_t)-\sum_{m=1}^{t-1} B(i_{m},m)  \right)^{\frac{1}{2}},\\
%\vspace{3 mm}
%B(i_j,t) = 0,\\
%B(i_j, t) = \frac{1}{B(i_t,t)}(K(i_j,i_t)  -\sum_{m=1}^{t-1}B(i_j,m)B(i_t,m)), \end{cases}\label{icd:algo}
%\end{align}
%for all $t<j\le n$.

%There are two possible choices of Cholesky decomposition. The incomplete Cholesky decomposition (ICD) stops its algorithm after $r$ steps and has complexity $O(nr^2)$. 
%To minimize the residue $\mathrm{tr}(K-BB^T)$, the $t$-th pivot is calculated at the end of the $t$-th interaction. 
%In the other hand, the complete Cholesky decomposition (CCD) consists of a full-rank decomposition, \textit{i.e.} the rank $r$ of $B$ is equal to $n$ and it has complexity $O(n^3)$. 
%In this case, the pivots are irrelevant to the final residue (which is zero) and can be assumed to be $i_t=t$. %The computation of pivots makes CCD more efficient for similar residues.
%We compare CCD and ICD performances in Fig. \ref{fig:residue}. We see that for small $r$, ICD is indeed faster. As we increase the value fo $r$, ICD becomes slower than CCD due to the pivot calculation but its residue decreases. We further compare both methods in Section \ref{time-scale}.

%We make sure $K$ is positive-definite by adding $\epsilon$ to its diagonal, where $\epsilon=\min(0,-\lambda_{min})$ and $\lambda_{min}$ is the smallest eigenvalue of $K$. Therefore, from the equation $BB^T=K+\epsilon\mathrm{Id}_n$, $B$ has rank $n$ and we can apply CCD.

%\subsection{Storage-efficient implementation}\label{low-rank} %/Low-rank decomposition} 
%Let us now consider instead the problem of improving storage efficiency. As discussed in section \ref{simi_score}, for each positive exemplar we store its original encoding plus a $2r$ vector, so we aim to decompose $K$ at a small rank $r$.
%For small values of $r$, a kernel PCA (KPCA) factorization guarantees a better approximation of $K$ than the Cholesky decomposition at a higher time cost. Indeed, as illustrated in Figure \ref{fig:residue}, a kernel PCA gives smaller residues than ICD at small rank.
